Palomar 5 Gaps

### 2024-10-28 paper/main.tex
trying to beef up motivation for this work

  One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way. ompared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo. For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information. \citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential. However, the use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time. For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, johnston05, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams. Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date based on a set of about 90 streams.

@@

  Our study focuses on the influence of globular clusters on stellar streams, specifically examining whether globular clusters could produce gaps similar to those expected from dark matter subhalos. If globular clusters do indeed create such gaps, this factor must be considered when using stellar streams to detect dark matter. An additional and unique aspect of our study is that we focus on modeling real stellar streams. We focus on the Palomar 5 stream, which is well-studied, has observable gaps \citep{2020ApJ...889...70B}, and is highly visible due to its position above the galactic disk. We present a result from one of our simulations in Fig.~\ref{fig:stream_on_sky}, and the following sections describe our modeling of the stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations affecting the stream.

>>>>

  One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way. Compared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo. For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information. \citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential. However, the use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time. For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, johnston05, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams. Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date based on a set of about 90 streams.

@@

  Previous studies have predominantly focused on the creation of gaps in stellar streams due to dark matter subhalo passages, with some exceptions that considered giant molecular clouds. If globular clusters can indeed create such gaps, this factor must be taken into account when using stellar streams to infer the population of dark matter subhaloes. In our study, we focus on Palomar 5, one of the most extensively studied and well-constrained globular cluster streams, known for its ease of observation. We present a result from one of our simulations in Fig. \ref{fig:stream_on_sky}. The following sections detail our modeling of the stream, its gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations impacting the stream.
